---
title: "GLM Assignment"
author: "Mike Swofford"
date: "11/9/2019"
output: html_document
# GLM Assignment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#options(scipen = 999)
```

```{r packages, echo=FALSE, include=FALSE}
library(tidyverse)
```



```{r data, echo=FALSE, include=FALSE}
# Import of dataset file
telework <- read_csv("Telework.csv")
```


```{r clean, echo=FALSE, include=FALSE}
# Cleaning of data

# Change state variable to a factor
cols <- c("state", "rent_or_own_home", "type_of_home", "education",
          "marital_status", "sex", "citizenship", "full_or_part_time", 
          "hourly_non_hourly", "union_member", "geography_region",
          "hispanic", "industry", "occupation_group", "detailed_industry", 
          "detailed_occupation_group", "metropolitan", "msa_size", "race",
          "occupation_code", "csa_code", "msa_code", "telecommute")
telework[cols] <- lapply(telework[cols], factor)
```

## Question 1: Telecommuting and Income

First, we will look at a model of weekly earnings (income) as it directly 
relates to whether someone telecommutes or not.

A simple one-way ANOVA provides these results:

```{r q1_anova, echo=FALSE}
# Question 1: ANOVA to measure weekly earnings as telecommute changes
q1 <- aov(weekly_earnings~as.factor(telecommute), data = telework)
q1_result <- summary(q1)
print(q1_result)
```
This model suggests that whether or not someone telecommutes has an impact on 
their income. The resulting p-value, being less than 2e-16, tells us that, given
the confidence level of 95%, there is good certainty that the dependent 
variable here, weekly earnings, is affected by the status of the dependent 
variable, the boolean telecommute.

A boxplot gives graphical represenation that the range of income for those that
telecommute is higher than the range of those that do not:

```{r q1_plot, echo = FALSE}

telework %>%
  ggplot(aes(y=weekly_earnings, x=(telecommute))) +
  geom_boxplot()
```


However, this should be considered a naive model for a couple of reasons:

1) Correlation between the variables should not imply causation. Logically, the
act of telecommuting is likely not in itself a driver of higher income. 
2) There may be other mediating variables in the data that give the appearance
of influence between these two selected variables. In this case, what is
more likely is that the type of occupation one has where telecommuting is an
option may drive higher income. 

## Question 2: Telecommuting, Income, and Full or Part-Time Status

```{r q2_anova, echo = FALSE}
# Question 2: Two-way ANOVA to measure weekly earnings as telecommute and 
# occupation type changes.

q2 <- aov(weekly_earnings~telecommute+full_or_part_time, 
          data = telework)
summary(q2)
TukeyHSD((q2))

telework %>%
  ggplot(aes(y=weekly_earnings, x=as.factor(full_or_part_time))) +
  geom_boxplot()


```


