---
title: "GLM Assignment"
author: "Mike Swofford"
date: "11/9/2019"
output: html_document
# GLM Assignment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE, include=FALSE}
library(car)
library(tidyverse)
```


```{r data, echo=FALSE, include=FALSE}
# Import of dataset file
telework <- read_csv("http://asayanalytics.com/telework_csv")
```

```{r clean, echo=FALSE, include=FALSE}
# Cleaning of data

# Change state of relevant variables to a factors
cols <- c("state", "rent_or_own_home", "type_of_home", "education",
          "marital_status", "sex", "citizenship", "full_or_part_time", 
          "hourly_non_hourly", "union_member", "geography_region",
          "hispanic", "industry", "occupation_group", "detailed_industry", 
          "detailed_occupation_group", "metropolitan", "msa_size", "race",
          "occupation_code", "csa_code", "msa_code", "telecommute")
telework[cols] <- lapply(telework[cols], factor)
```

## Question 1: Telecommuting and Income

First, we will look at a model of weekly earnings (income) as it directly 
relates to whether someone telecommutes or not. 

A simple one-way ANOVA provides these results:

```{r q1_anova, echo=FALSE}
# Question 1: ANOVA to measure weekly earnings as telecommute changes

q1 <- aov(weekly_earnings~telecommute, data = telework)
q1_result <- summary(q1)
print(q1_result)
```
A. This model suggests that whether or not someone telecommutes has an impact on 
their income. The resulting p-value, being less than 2e-16, tells us that, given
the confidence level of 95%, there is good certainty that the dependent 
variable here, weekly earnings, is affected by the status of the dependent 
variable, the boolean telecommute variable.

B. A boxplot gives graphical represenation that the range of income for those 
that telecommute is higher than the range of those that do not:

```{r q1_plot, echo = FALSE}
# Boxplot of weekly_earning, telecommute

telework %>%
  ggplot(aes(y=weekly_earnings, x=telecommute)) +
  geom_boxplot()
```


C. This should be considered a naive model for a couple of reasons:

1) Correlation between the variables should not imply causation. Logically, the
act of telecommuting is likely not in itself a driver of higher income. 
2) There may be other mediating variables in the data that give the appearance
of influence between these two selected variables. In this case, what is
more likely is that the type of occupation one has where telecommuting is an
option may drive higher income. 

---

## Question 2: Telecommuting, Income, and Full or Part-Time Status

A second model includes full or part-time status as a variable in addition to 
the telecommute status.

A two-way ANOVA gives these results:

```{r q2_anova, echo = FALSE}
# Question 2: Two-way ANOVA to measure weekly earnings as telecommute and 
# occupation type changes.

q2 <- aov(weekly_earnings~telecommute+full_or_part_time, 
          data = telework)
summary(q2)
```

A. There is logic to this model, as one would assume that holding full-time 
employment would have a direct influence on an increase in income, while a 
part-time worker would likely have a compartively lower income. 


B. In this model, both variables show significance, as the p-values in the above 
results are both very low. I.e. both telecommute status and full or part-time 
status show a correlation to income. 


Concerning the telecommute status, however, there may still be some correllation
to income, but this is likely mediated by the full-time/part-time status
variable. If that is the case, then this model still shows some naievety as
there is multicollinearity between the variables.

C. A boxplot of this model shows the similarities in the range variancess for
those that telecommute and those that do not. While weekly earning are higher
for full-time workers that telecommute, the similar dynamic between the ranges
of the two telecommute categories is apparent:

```{r q2_plot, echo = FALSE}
# Plot of weekly_earnings and telecommute, sepearted by full/part time status.

telework %>%
  ggplot(aes(y=weekly_earnings, x=telecommute)) +
  geom_boxplot() + aes(fill = full_or_part_time)
```

D. Comparing the first model with only telecommute as a variable to this model
that includes full or part time status, this new model appears to be a better 
fit:

```{r q2_anova_comp, echo = FALSE}
# Comparison of ANOVAs from question 1 and question 2

q1_anova <- aov(weekly_earnings~telecommute, data = telework)
summary(q1_anova)
summary(q2)
```
The comparison between these ANOVA results shows that the low p-values and the
sum of squares indicates a greater fit with the additional variable.


---

## Question 3: Simple Regression of Income / Hours Worked

Next is a simple linear regression model for weekly earnings (income) as a 
function of the number of hours worked.

A. The general equation for this model is: $$\hat Y = b_{0} + b_{1}X_{i}$$

B. Incorporating the above variables, the specific form for this model will be:
$$weekly~earnings_{i} = \beta_{0} + \beta_{1}(hours~worked)$$
The results of the regression are:

```{r q3_lm, echo = FALSE}
# Linear model of weekly_earnings and hours_worked

summary(q3_result <- lm(weekly_earnings~hours_worked, data = telework))
```

From this, the specific equation for the model is:

$$weekly~earnings_{i} = {`r (as.numeric(coef(q3_result)[1]))`} + 
{`r (as.numeric(coef(q3_result)[2]))`}(hours~ worked)$$
C. This is a naive model for several reasons:
<ol>
<li>While hours worked may have influence on weekly earnings given comparable
circumstances (same job, same industry), hours worked alone does not have
significant influence on income.</li>
<li>Hours worked is also likely a function of full-time or part-time status, as
one who works full-time is also likely to have a higher number of hours worked.
</li>
<li>The affect of hourly or non-hourly status also likely has a strong influence
on hours worked. Non-hourly (salaried) employees are typically in positions of 
management or professional occupations. They may work near 40 hours a week, but
make significantly higher income.</li>
</ol>

D. Using only these two variables, this is not a helpful model. The above stated
affect of hourly and non-hourly employment certainly should be considered. We 
should expect that the hours worked would have a greater influence on hourly 
workers than non-hourly.

With only the two variables of income and weekly earnings, the model appears as:

```{r q3_plot, echo = FALSE}
# Plot of weekly_earnings and hours_worked linear model

telework %>%
  ggplot(aes(x=hours_worked, y=weekly_earnings))+
 geom_point()+
 geom_smooth(method="lm")
```

E. However, when accounting for the hourly/non-hourly status, the models show
that hours worked appears to have a more correlated impact on weekly earnings.
In the graphs below, the slimmer scope of the shaded area of the left 
visualization (hourly) as compared to the right (non-hourly) shows that these 
categories are affected differently by the hourly/non-hourly status.

```{r q3_adjusted, echo = FALSE}
# Plot of weekly_earnings and hours worked linear model, shown by 
# hourly_non_hourly status.

telework %>%
    ggplot(aes(x=hours_worked, y=weekly_earnings, color = hourly_non_hourly))+
    geom_point()+
    facet_wrap(telework$hourly_non_hourly) +
    geom_smooth(method="lm")
```

---

## Question 4: Simple Regression of Income / Age

Next is a simple linear regression model for weekly earnings (income) as a 
function of age.

A. The general equation for the model is: $$\hat Y = b_{0} + b_{1}X_{i}$$

B. Incorporating the above variables, the specific form for this model will be:
$$weekly~earnings_{i} = \beta_{0} + \beta_{1}(age)$$



```{r q4_lm, echo = FALSE}
# Linear model of weekly_earnings and age

q4_result <- lm(weekly_earnings~age, data = telework)
```
These results provide estimated form of the regression as:
$$weekly~earnings_{i} = 548.9457 + 9.1941(age)$$

C. This model is naive for several reasons:
<ol>
<li>This model may account for the increasing income an employee would 
experience with age, but it ignores factors of how much work is done such as
part-time/full-time status, hours worked, and hourly/non-hourly status.</li>
<li>There are likely large influences in earning based on geographic location, 
e.g. a 50-year-old in California with the same occupation as a 50-yearl-old in 
Wyoming likely has much higher earnings.</li>
<li>Other demographic information such as sex and citizenship are not accounted
for in this model.</li>
</ol>

D. Testing the linearity assumptions of this model can start with the summary of
the linear model:

```{r q4_summary, echo = FALSE}
# Show summary stats of linear model

summary(q4_result)
```

These results show deficiences with the model:
<ol>
<li>The R-Squared is very low at .03696, suggesting the model is not a good fit.
</li>
<li>The residual standard error is significant. At 654.6, this shows a wide
variability in the residuals, likely caused by a number of outliers.</li>
</ol>

Another test is to visualize the residuals against the fitted values: 

```{r q4_plot, echo = FALSE}
# Residual plot of linear model

plot(q4_result, 1)
```


From this plot, the curve of the line shows there is not a good fit betwen the
fitted values and the residuals.

E. Other concerns become apparent looking at the plot of this model:

```{r q4_visualization, echo = FALSE}
# Plot of weekly_earnings and age linear model

telework %>%
  ggplot(aes(x=age, y=weekly_earnings)) +
    geom_point()+
    geom_smooth(method="lm")
```


From this graph, several concerns are apparent:
<ol>
<li>There appears to be a cap or placeholder value for weekly earnings at 
around $2800. Surely some employees make more than this amount. The use of a
placeholder or a cap is skewing the data towards that value. Data should be 
collected for all income levels, or this placeholder should be removed and 
accounted for in a model</li>
<li>There is no data between the ages of 80 and about 85. Every other age is
represented. The data is therefore skewed away from this range. If data is not
available for this range, the best solution would be to remove or filter out
the data for employees at age 85 and qualify the data as only up to 80.</li>
</ol>

---

## Question 5: Linearity with Multiple Variables

This final model will be a regression of weekly earnings with the following 
variables:
<ul>
<li>X <sub>1</sub> = age</li>
<li>X <sub>2</sub> = sex</li>
<li>X <sub>3</sub> = full-time/part-time status</li>
<li>X <sub>4</sub> = hourly/non-hourly status</li>
</ul>

A. The regression equation is:
$$\hat Y = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + \beta_4X_4$$
```{r q5_lm, echo = FALSE}
# Linear model of weekly_earnings, age, sex, full_or_part_time and 
# hourly_or_non_hourly

q5_result <- lm(weekly_earnings~age + sex + full_or_part_time +
                  hourly_non_hourly, data = telework)
```

The results for this model are:

```{r q5_result, echo = FALSE}
# Summary stats of linear model

summary(q5_result)
```

B. From the results of the regression, the model becomes:

$$weekly\ earnings = 611.7809 + 6.6874(age) - 179.994(sex) - 
493.1572(full/part\ status) + 479.0939(hourly/nonhourly\ status)$$

C. There does not appear to be any obvious colinearity, as 
shown by a variance inflation factors (vif) test where all the
variables test with a low result:

```{r q5_vif, echo = FALSE}
# Varirance inflaction factors of linear model to test colinearity

vif(q5_result)
```
D. The best ranges to use for the variables can be computed using the confidence
intervals from the linear model:

```{r q5_confint, echo = FALSE}
# Compute confidence intervals for linear model

ranges <- confint(q5_result)
print(ranges)
```

E. A hypothetical observation might be a person who is 40 years old, male, works
full-time, and is non-hourly. Using this model, the equation for this
observation would be:

$$611.7809 + 6.6874* age + -179.9994 * sex + -493.1572 * full\_or\_part\_time +
479.0939 * hourly\_non\_hourly$$

```{r q5_predict, echo = FALSE}
# Define hypothetical observation variables

age <- 40
sex <- 1
time <- 2
hourly <- 2

# Calculate result of hypothetical observation
result <- round(611.7809 + 6.6874*(age) + -179.9994 *(sex) + -493.1572 * time + 
                  479.0939 * hourly, 2)

# Calculate results of hypothetical observation using confidence intervals
upper <- round(ranges[1, 1] + (ranges[2, 1]* age) + (ranges[3, 1] * sex) + 
  (ranges[4, 1] * time) + (ranges[5, 1] * hourly), 2)

lower <- round(ranges[1, 2] + (ranges[2, 2]* age) + (ranges[3, 2] * sex) + 
  (ranges[4, 2] * time) + (ranges[5, 2] * hourly), 2)
```

From this model, the predicted weekly earnings for this observation would be 
**$`r result`**.

Using the ranges above, the weekly earnings lower limit would be **$`r lower`**
and the upper limit would be **$`r upper`**.